---
title: "Task 2.3"
author: "Aj Averett"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
 html_document:
  keep_md: TRUE
  toc: true
  toc_float: true
  code_folding: show
---

# Setup

```{r}
#install.packages("gapminder") #uncomment and run once. 
library(gapminder)
library(tidyverse)
```
# The Gapminder Data

After watching Hans Rosling’s original TED talk, some insights include - My agreement in that public data should be made accessible and easy to use
- Data from virtually any domain can be made understandable through the expanse of human boredom and curiosity.

## Pick a quantitative variable and create a chart that summarizes the distribution.


```{r}
temp <- gapminder[gapminder$country == 'United States',]

gapminder[gapminder$country == 'United States',] %>%
  ggplot(aes(x=year, y=lifeExp)) +
    geom_line() +
    theme_classic()


```


## Pick a qualitative variable and create a chart that summarizes the distribution.


```{r}
gapminder[gapminder$year == 2007,] %>% 
  ggplot(aes(x = continent, fill = continent)) +
  geom_bar() +
  theme_classic

```

## Pick two variables and create a chart that summarizes the bivariate distribution (the relationship between the two).


```{r}
gapminder[gapminder$year == 2007,] %>% 
  ggplot(aes(x = lifeExp, y = gdpPercap, color = continent)) + 
    geom_point() + 
  theme_classic() +
  scale_y_log10()
  
```

## Try using color, shape, alpha, facets, etc., to visualize the relationships among 3 or more variables.


```{r}

gapminder[gapminder$year == 2007,] %>%
  ggplot(aes(x = reorder(continent,lifeExp), y = lifeExp, fill = continent)) +
    geom_boxplot() +
    theme_classic()

```

## Use the rest of your preparation time for this task to repeat the above on various variables.


```{r}
gapminder[gapminder$year == 2007,] %>%
  ggplot(aes(x=gdpPercap, group=continent, fill=continent)) +
    geom_density(adjust=1.5, alpha=.3) +
    theme_classic()
```




  