---
title: "Project_1"
author: "Aj Averett"
date: '2022-04-22'
output:
  rmdformats::readthedown:
    code_folding: hide
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background Data

Throughout this project, we will analyze the brightness of LED light bulbs over time. Interestingly, LED light bulbs will increase in brightness before their steady decrease and ultimate burnout. This process may last up to three years. Since it is not feasible to test the bulb lifetime- this project will utilize mathematical models to predict the lifespan of some of these bulbs.

## Task One

In this R Markdown file, the led_bulb function will be used to generate random bulbs for predictions.

```{r}
library(data4led)
bulb <- led_bulb(1,seed = 196)
```

The "head" function below verifies that the data frame, bulb, includes the columns: id, hours, intensity, demonormalized_intensity, and percent_intensity.


```{r}
head(bulb)
```

The code below creates a plot from the bulb data frame. The plot has hours on the horizontal axis (x-axis) and percent_intensity on the vertical axis (y-axis).

```{r}
plot(percent_intensity ~ hours, data = bulb)
```

